\en{%Let $n\geq 2$ be an integer. Switzerland and Liechtenstein are performing their annual festive show. There is a field subdivided into $n \times n$ squares, in which the bottom-left square contains a red house with $k$ Swiss gymnasts, and the top-right square contains a blue house with $k$ Liechtensteiner gymnasts. Every other square only has enough space for a single gymnast at a time. Swiss gymnasts can only move to either the square immediately above or to the right, and Liechtensteiner gymnasts can only move either to the square immediately below or to the left. The goal is to move all the Swiss gymnasts to the blue house and all the Liechtensteiner gymnasts to the red house, with the caveat that a gymnast of either nationality cannot enter a house until all the gymnasts of the other nationality have left. Depending on $n$, find the largest $k$ for which this is possible.

Let $n\geq 2$ be an integer. Switzerland and Liechtenstein are performing their annual festive show. There is a field subdivided into $n \times n$ squares, in which the bottom-left square contains a red house with $k$ Swiss gymnasts, and the top-right square contains a blue house with $k$ Liechtensteiner gymnasts. Every other square only has enough space for a single gymnast at a time. Each second either a Swiss gymnast or a Liechtensteiner gymnast moves. The Swiss gymnasts moves to either the square immediately above or to the right and the Liechtensteiner gymnasts moves either to the square immediately below or to the left. The goal is to move all the Swiss gymnasts to the blue house and all the Liechtensteiner gymnasts to the red house, with the caveat that a gymnast cannot enter a house until all the gymnasts of the other nationality have left. Determine the largest $k$ in terms of $n$ for which this is possible.}

\de{Sei $n\geq 2$ eine ganze Zahl. Die Schweiz und Liechtenstein führen ihre alljährliche Festvorstellung auf. Die Bühne ist in $n \times n$ Quadrate unterteilt, wobei das untere linke Quadrat ein rotes Haus mit $k$ Schweizer Turnern enthält und das obere rechte Quadrat ein blaues Haus mit $k$ Liechtensteiner Turnern. Jedes andere Quadrat bietet jeweils nur Platz für einen einzigen Turner. Jede Sekunde bewegt sich entweder ein Schweizer Turner oder ein Liechtensteiner Turner von seinem Quadrat auf ein direkt angrenzendes Quadrat. Schweizer Turner bewegen sich stets nach oben oder nach rechts und Liechtensteiner Turner bewegen sich stets nach unten oder nach links. Ziel ist es, alle Schweizer Turner in das blaue Haus und alle Liechtensteiner Turner in das rote Haus zu bringen, wobei ein Turner erst dann ein Haus betreten darf, wenn alle Turner der anderen Nationalität das Haus verlassen haben. Bestimme das grösste $k$ abhängig von $n$, für welches dies möglich ist.}

\fr{Soit $n\geq 2$ un nombre entier. La Suisse et le Liechtenstein présentent leur spectacle festif annuel. Il y a un champ subdivisé en $n\times n$ carrés. Le carré en bas à gauche contient une maison rouge avec $k$ gymnastes suisses, et le carré en haut à droite contient une maison bleue avec $k$ gymnastes liechtensteinois. Chaque autre carré a seulement la place pour un seul gymnaste à la fois. Chaque seconde, soit un gymnaste suisse, soit un gymnaste liechtensteinois se déplace. Les gymnastes suisses se déplacent en allant sur le carré au-dessus ou le carré à droite et les gymnastes liechtensteinois se déplacent en allant soit sur le carré au-dessous ou le carré à gauche. Le but est que tous les gymnastes suisses se déplacent jusqu'à la maison bleue, et que tous les gymnastes liechtensteinois se déplacent jusqu'à la maison rouge, tout en sachant qu'un gymnaste ne peut pas entrer dans une maison avant que tous les gymnastes de l'autre nationalité l'aient quittée. Déterminer la valeur maximale de $k$ en fonction de $n$ pour laquelle cela est possible.}