\en {
Let $(m, n)$ be a pair of positive integers. Julia has carefully planted $m$ rows of $n$ dandelions in an $m \times n$ array in her back garden. Now, Jana and Viviane decide to play a game with a lawnmower they just found. Taking alternating turns and starting with Jana, they can mow down all the dandelions in a straight horizontal or vertical line (and they must mow down at least one dandelion!). The winner is the player who mows down the final dandelion. Determine all pairs $(m,n)$ for which Jana has a winning strategy.
}

\de {
Sei $(m, n)$ ein Paar von natürlichen Zahlen. Julia hat sorgfältig $m$ Reihen mit jeweils $n$ Löwenzahn in einem $m \times n$ Raster in ihrem Garten gepflanzt. Nun spielen Jana und Viviane ein Spiel mit einem Rasenmäher, den sie gerade gefunden haben. Jana beginnt, und nacheinander können sie jeweils alle Löwenzahn in einer geraden horizontalen oder vertikalen Linie mähen (und sie müssen mindestens einen Löwenzahn mähen!). Diejenige, die den letzten Löwenzahn mäht, gewinnt das Spiel. Bestimme alle Paare $(m,n)$, für welche Jana eine Gewinnstrategie hat.
}

\fr {Soit $(m,n)$ une paire d'entiers strictement positifs. Julia a scrupuleusement planté $m$ rangées de $n$ pissenlits dans son jardin de telle manière que les pissenlits forment une grille de taille $m\times n$. Jana et Viviane décident de jouer à un jeu avec une tondeuse qu'elles viennent de trouver. L'une après l'autre, en commençant par Jana, elles tondent tous les pissenlits dans une rangée ou dans une colonne (et elles doivent toujours tondre au moins un pissenlit). La gagnante est celle qui tond le dernier pissenlit. Déterminer toutes les paires $(m,n)$ pour lesquelles Jana a une stratégie gagnante.

}

\ita {

}