\en{
Let $n$ be a positive integer. The islands of the MO-Archipelago are arranged in a regular equilateral triangular unitary grid to form a big equilateral triangle of side length $n$. The beloved Governor Henning is in charge of building bridges between every pair of islands that are a distance of $1$ apart. For every island $i$, Henning chooses two real numbers $x_i$ and $y_i$ that satisfy $x_i^2+y_i^2 = 1$. The cost of a bridge between islands $i$ and $j$ is then given by $1+ x_i\, x_j + y_i\, y_j$. Determine the minimal amount of money needed to build all the bridges.

\bigskip

\textbf{Answer:} The minimal cost is $\frac{3n(n+1)}{4}$.

\bigskip

\textbf{Solution (Henning):} Let us start with the observation that one can choose $\frac{n (n+1)}{2}$ unit triangles such that each bridge is contained exactly once in one of the triangles. For each triangle one then has the following inequality:
\begin{align*} 
(x_1 + x_2 +x_3)^2 + (y_1+y_2 +y_3)^2  &\geq 0  \\
6+2(x_1 x_2 + x_2 x_3+ x_3x_1) +2(y_1 y_2 + y_2 y_3+ y_3y_1) &\geq -(x_1^2+x_2^2 + x_3^2)- (y_1^2 +y_2^2 + y_3^2) +6 = 3  
\end{align*}
where $x_i, y_i$ belong to the three islands in the unit triangle. Summing now over all unit triangles we arrive at the lower bound of $$\frac{3n(n+1)}{4}$$ as the cost for all the bridges. \\

We now show that this lower bound can be reached. One can find that the equality case for one of the unit triangles is fulfilled by choosing 

\[ ((x_1,y_1), (x_2, y_2), (x_3, y_3)) = \left( (1,0),(-\dfrac{1}{2}, \dfrac{\sqrt{3}}{2}),(- \dfrac{1}{2},-\dfrac{\sqrt{3}}{2}) \right)  \]

It is now important to remark that the equality case can be achieved setting each of the $(x_i, y_i)$ in the big triangle to one of three above vectors. Calling the first vector the type 1 vector, the second and third vector type 2 and 3 respectively, the problem turns into a colouring exercise. \\

If we start to fill the lowest row of the big triangle according to the $1, 2, 3, 1, 2... $ pattern. The row above with $3, 1, 2, 3, 1 ..$ and so on, we see that the islands are now coloured in a way that each of the chosen unit triangles has exactly one island of each type, hence the equality case is fulfilled for all of the unit triangles simultaneously.  


\textbf{Marking Scheme}
\begin{enumerate}
\item 2P: Finding the minimum for one unit triangle. 
\item 1P: Showing that the equality case can be achieved for a unit triangle. 
\item 1P: Marking the upside unit triangles. 
\item 1P: Realising that any construction or lower bound on an upside unit triangle can easily generalised to the whole problem.
\item 2P: Showing that the equality case cam be achieved for all unit triangles simultaneously 
\end{enumerate}
}