\en{Let $p$ be an odd prime. Arnaud has hung up $N\geq 1$ towels to dry on a washing line, each coloured either purple or yellow. Then, for each $1 \leq n \leq N$, he calculates what fraction of the first $n$ towels are yellow, and writes down these $N$ fractions in their irreducible form on a piece of paper. Julia finds the piece of paper the next day and notices that all of the fractions $\frac{1}{p}, \frac{2}{p}, \dots, \frac{p-1}{p}$ are on the paper. Prove that 
\[
N \geq \frac{p^3-p}{4}.
\]}


\de{Sei $p$ eine ungerade Primzahl. Arnaud hat $N\geq 1$ Tücher an einer Wäscheleine aufgehängt, wobei jedes Tuch entweder violett oder gelb ist. Nun berechnet er für jedes $1 \leq n \leq N$, welcher Bruchteil der ersten $n$ Tücher gelb ist, und notiert diese $N$ Brüche in gekürzter Form auf einem Blatt Papier. Julia findet das Blatt Papier am nächsten Tag und stellt fest, dass alle Brüche $\frac{1}{p}, \frac{2}{p}, \dots, \frac{p-1}{p}$ darauf vorkommen. Beweise, dass
\[
N \geq \frac{p^3-p}{4}.
\]}

\fr{Soit $p$ un nombre premier impair. Arnaud a suspendu $N\geq 1$ tee-shirts sur une corde à linge. Chaque tee-shirt est soit violet, soit jaune. Il calcule ensuite, pour chaque $1\leq n\leq N$, la fraction des $n$ premiers tee-shirts qui sont jaunes et écrit ces $N$ fractions sous forme irréductible sur une feuille. Julia trouve la feuille le lendemain et constate que les fractions $\frac{1}{p}, \frac{2}{p}, \dots, \frac{p-1}{p}$ apparaissent toutes sur la feuille. Montrer que 
\[
N \geq \frac{p^3-p}{4}.
\]
}